Local PCA Shows How the Effect of Population Structure Differs Along the Genome

### 2018-08-22 writeup/review-responses.tex

\begin{point}{PC switching}
\ldots could lead to a small number of such windows having their
orientation reversed: that is, PC1 becomes PC2 and vice versa. (Or PC2 and PC3
could switch).
\end{point}

\reply{
    This is a natural concern.
    However, the only point at which we compare PCs in a way that could be sensitive to ordering
    is in determining the window size -- in computing the distance between windows
    we use a measure which is invariant under ordering.
    We have made this more clear by moving the note about flipping signs of PCs
    to the appendix on window choice \llname{ll:moved_pc_note}
    and added more explicit notes about this to \llname{ll:another_pc_point} and \revref.
    Furthermore, one of our simulation tests
    is explicitly designed to test for this effect. \llname{ll:pc_switching_test}
}

\begin{point}{p6}
``here, we use k=2...''  - you have to show that $k>2$ is the same.
\end{point}

\reply{
    We have added a table including correlations between MDS values between $k=2$ and $k=5$
    for the Medicago data (Table \ref{tab:param_cors}),
    and a representative figure (for the Drosophila data, $k=5$),
    see Figure \ref{fig:dmel_k5}. \revref
}

@@

\begin{point}{p15}
``We also found nearly identical results when choosing shorter windows of 1,000 SNPs'' - again, show this.
\end{point}

\reply{
    The same table includes correlations between MDS values for different window lengths 
    (Table \ref{tab:param_cors}). \revref
}

@@

\begin{point}{p15}
 ``or choosing windows of equal length in base pairs rather than SNPs'' - once again.
\end{point}

\reply{
    Again, we address this in Table \ref{tab:param_cors}. \revref
}

@@

\begin{point}{}
``There has been substantial debate over the relative impacts of different forms of selection.''
Citation needed.
\end{point}

\reply{
    We have added some relevant recent citations,
    but apologize for not being aware of the appropriate recent review to refer to.
    \revref
}

\begin{point}{}
``Results using larger numbers of PCs were nearly identical''. It would be interesting to have a
supplementary table.
\end{point}

\reply{
    We have included a table of correlations (Table \ref{tab:param_cors}),
    as well as the plots for Drosphila data using $k=5$
    (Figure \ref{fig:dmel_k5}),
    and hope that readers will take our word on this point more generally.
}

### 2018-08-22 writeup/local_pca_paper.tex
better

We also found nearly identical results when choosing shorter windows of 1,000 SNPs
(Supplemental Figure \ref{sfig:mds_medicago_allchr_1e3}; \revpoint{1}{4}
or choosing windows of equal length in base pairs rather than SNPs. \revpoint{1}{5}
% Similarly, the results were not substantially changed
% when using weighted PCA to downweight the large group of Tunisian samples.

@@

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Wed Feb  8 16:17:12 2017
\begin{table}[ht]
\centering
    \begin{tabular}{rrrrrr}
      \hline
          & 10000 SNPs & 1000 SNPs & 10000 SNPs & 100000bp & 10000bp \\ 
     MDS1 & 2 PCs & 2 PCs & 5 PCs & 2 PCs & 2 PCs \\ 
      \hline
  10000 SNPs, 2 PCs & 1.00 & 0.87 & 0.96 & 0.90 & 0.88 \\ 
  1000 SNPs, 2 PCs & 0.68 & 1.00 & 0.73 & 0.68 & 0.94 \\ 
  10000 SNPs, 5 PCs & 0.96 & 0.92 & 1.00 & 0.88 & 0.93 \\ 
  100000bp, 2 PCs & 0.90 & 0.87 & 0.88 & 1.00 & 0.87 \\ 
  10000bp, 2 PCs & 0.68 & 0.93 & 0.72 & 0.67 & 1.00 \\ 
       \hline
       \hline
        MDS2 & &&&&\\
      \hline
      10000 SNPs, 2 PCs & 1.00 & 0.54 & 0.93 & 0.87 & 0.56 \\ 
      1000 SNPs, 2 PCs & 0.82 & 1.00 & 0.76 & 0.83 & 0.92 \\ 
      10000 SNPs, 5 PCs & 0.93 & 0.50 & 1.00 & 0.83 & 0.52 \\ 
      100000bp, 2 PCs & 0.87 & 0.59 & 0.84 & 1.00 & 0.58 \\ 
      10000bp, 2 PCs & 0.83 & 0.92 & 0.77 & 0.84 & 1.00 \\ 
       \hline
    \end{tabular}
    \begin{caption}
        Correlations between MDS coordinates of genomic regions
        between runs with different parameter values.
        To produce these, 
        we first ran the algorithm with the specified window size and number of PCs 
        ($k$ in equation \plr{XXX})
        on the full \textit{Medicago truncatula} dataset.
        Then to obtain the correlation between results obtained from parameters A in the row of the matrix above
        and parameters B in the column of the matrix above,
        we mapped the windows of B to those of A by averaging MDS coordinates of any windows of B whose midpoints lay in the corresponding window of A;
        we then computed the correlation between the MDS coordinates of A and the averaged MDS coordinates of B.
        This is not a symmetric operation, so these matrices are not symmetric.
        As expected, parameter values with smaller windows produce noisier estimates.
        Plots of MDS values along the genome are visually nearly identical
        for parameter sets having similar window sizes --
        full reports are available as supplementary material on Data Dryad \plr{XXX}.
        \label{tab:param_cors}
    \end{caption}
\end{table}

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=1\textwidth]{Fig1_allchr_Together_MDS_plot_PCAk5_win103}
    \end{center}
    \caption{      
        MSD visualizations for each chromosome arm of \textit{Drosophila melanogaster},
        as in \Figure \ref{fig:mds_allchr},
        except that the method was run using five PCs ($k=5$) instead of two.
         \label{fig:dmel_k5}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
        \ifusepngs
           \includegraphics[width=0.85\textwidth]{Medicago_MDS_plot_allchr_win103_PCAk2.png}
       \else
           \includegraphics[width=0.85\textwidth]{Medicago_MDS_plot_allchr_win103_PCAk2}
       \fi
    \end{center}
    \caption{
        MDS visualizations of the effects of population structure for all 8 chromosomes 
        of the \textit{Medicago truncatula} data as in \Figure \ref{sfig:mds_medicago_allchr},
        but using windows of $10^3$ SNPs.
        \label{sfig:mds_medicago_allchr_1e3}
    }
\end{figure}

>>>>

The results were highly consistent across window sizes, window types (SNPs or bp),
and number of PCs (Table \ref{tab:param_cors}).
\revpoint{1}{4} \revpoint{1}{5} \revpoint{1}{6}
% We also found nearly identical results when choosing shorter windows of 1,000 SNPs
% (Supplemental Figure \ref{sfig:mds_medicago_allchr_1e3}; \revpoint{1}{4}
% or choosing windows of equal length in base pairs rather than SNPs. \revpoint{1}{5}
% Similarly, the results were not substantially changed
% when using weighted PCA to downweight the large group of Tunisian samples.

@@

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Wed Feb  8 16:17:12 2017
\begin{table}[ht]
\centering
    \begin{tabular}{rrrrrr}
      \hline
          & 10000 SNPs & 1000 SNPs & 10000 SNPs & 100000bp & 10000bp \\ 
     MDS1 & 2 PCs & 2 PCs & 5 PCs & 2 PCs & 2 PCs \\ 
      \hline
  10000 SNPs, 2 PCs & 1.00 & 0.87 & 0.96 & 0.90 & 0.88 \\ 
  1000 SNPs, 2 PCs & 0.68 & 1.00 & 0.73 & 0.68 & 0.94 \\ 
  10000 SNPs, 5 PCs & 0.96 & 0.92 & 1.00 & 0.88 & 0.93 \\ 
  100000bp, 2 PCs & 0.90 & 0.87 & 0.88 & 1.00 & 0.87 \\ 
  10000bp, 2 PCs & 0.68 & 0.93 & 0.72 & 0.67 & 1.00 \\ 
       \hline
       \hline
        MDS2 & &&&&\\
      \hline
      10000 SNPs, 2 PCs & 1.00 & 0.54 & 0.93 & 0.87 & 0.56 \\ 
      1000 SNPs, 2 PCs & 0.82 & 1.00 & 0.76 & 0.83 & 0.92 \\ 
      10000 SNPs, 5 PCs & 0.93 & 0.50 & 1.00 & 0.83 & 0.52 \\ 
      100000bp, 2 PCs & 0.87 & 0.59 & 0.84 & 1.00 & 0.58 \\ 
      10000bp, 2 PCs & 0.83 & 0.92 & 0.77 & 0.84 & 1.00 \\ 
       \hline
    \end{tabular}
    \caption{
        Correlations between MDS coordinates of genomic regions
        between runs with different parameter values.
        To produce these, 
        we first ran the algorithm with the specified window size and number of PCs 
        ($k$ in equation \eqref{eqn:est_cov})
        on the full \textit{Medicago truncatula} dataset.
        Then to obtain the correlation between results obtained from parameters A in the row of the matrix above
        and parameters B in the column of the matrix above,
        we mapped the windows of B to those of A by averaging MDS coordinates of any windows of B whose midpoints lay in the corresponding window of A;
        we then computed the correlation between the MDS coordinates of A and the averaged MDS coordinates of B.
        This is not a symmetric operation, so these matrices are not symmetric.
        As expected, parameter values with smaller windows produce noisier estimates,
        but plots of MDS values along the genome are visually very similar.
    } \label{tab:param_cors}
\end{table}

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=1\textwidth]{Fig1_allchr_Together_MDS_plot_PCAk5_win103}
    \end{center}
    \caption{      
        MDS visualizations for each chromosome arm of \textit{Drosophila melanogaster},
        as in \Figure \ref{fig:mds_allchr},
        except that the method was run using five PCs ($k=5$) instead of two.
         \label{fig:dmel_k5}
    }
\end{figure}

@@

% \begin{figure}
%     \begin{center}
%         \ifusepngs
%            \includegraphics[width=0.85\textwidth]{Medicago_MDS_plot_allchr_win103_PCAk2.png}
%        \else
%            \includegraphics[width=0.85\textwidth]{Medicago_MDS_plot_allchr_win103_PCAk2}
%        \fi
%     \end{center}
%     \caption{
%         MDS visualizations of the effects of population structure for all 8 chromosomes 
%         of the \textit{Medicago truncatula} data as in \Figure \ref{sfig:mds_medicago_allchr},
%         but using windows of $10^3$ SNPs.
%         \label{sfig:mds_medicago_allchr_1e3}
%     }
% \end{figure}
